A comparative risk assesment of U.S. regional active transport using the ITHIM model

### 2016-10-05 R/ITHIM.Rmd

```{r compute, echo = FALSE, eval = TRUE}
compute <- FALSE
```

@@

```{r subset, eval = compute, fig.width = 10, echo = FALSE}
NHTS.person <- subset( NHTS.person, MSA == TRUE & HHSTATE %in% c("CA","TX","NY","FL","VA","NC","AZ"))
NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
saveRDS(NHTS, file = "./data/NHTS.rds")
```

@@

```{r roadInjuries, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
roadFatalaties <- read.csv(file = "./../data/BikePedDeaths_09-13.csv", na.strings = c("998","999"))
roadFatalaties <- within(roadFatalaties, {
  ageClass <- convertToAgeClass(AGE)
  SEX <- ifelse(SEX == 1, "M", ifelse(SEX == 2, "F", NA))
  SEX <- factor(SEX, levels = c("F","M"))
})
roadFatalaties <- subset(roadFatalaties, !is.na(SEX))

p <- ggplot(roadFatalaties, aes(ageClass, fill = SEX))
p + geom_bar(position = "dodge") + theme_bw()+ theme(axis.text.x  = element_text(angle=90, vjust=0.5)) + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + facet_grid( death ~ ., scales = "free_y")
```

>>>>

```{r compute, echo = FALSE, eval = TRUE}
compute <- TRUE
```

@@

```{r subset, eval = compute, fig.width = 10, echo = FALSE}
NHTS.person <- subset( NHTS.person, MSA == TRUE ) # & HHSTATE %in% c("CA","TX","NY","FL","VA","NC","AZ"))
NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
saveRDS(NHTS, file = "./data/NHTS.rds")
```

@@

```{r roadInjuries, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
convertState <- function( stateCode ){

stateCode <- as.character(stateCode)

stateCodeMatrix <- matrix(c("Alabama","1","Nebraska","31","Alaska","2","Nevada","32","Arizona","4","New Hampshire","33","Arkansas","5","New Jersey","34","California","6","New Mexico","35","Colorado","8","New York","36","Connecticut","9","North Carolina","37","Delaware","10","North Dakota","38","District of Columbia","11", "Ohio", "39","Florida","12","Oklahoma","40","Georgia","13","Oregon","41","Hawaii","15","Pennsylvania","42","Idaho","16","Puerto Rico","43","Illinois","17","Rhode Island","44","Indiana","18","South Carolina","45","Iowa","19","South Dakota","46","Kansas","20","Tennessee","47","Kentucky","21","Texas","48","Louisiana","22","Utah","49","Maine","23","Vermont","50","Maryland","24","Virgin Islands (since 2004)","52","Massachusetts","25","Virginia","51","Michigan","26","Washington","53","Minnesota","27","West Virginia","54","Mississippi","28","Wisconsin","55","Missouri","29","Wyoming","56","Montana","30"), byrow = TRUE, ncol = 2)

stateAbbrevMatrix <- matrix(c("Alabama","AL","Alaska","AK","Arizona","AZ","Arkansas","AR","California","CA","Colorado","CO","Connecticut","CT","Delaware","DE","District of Columbia","DC","Florida","FL","Georgia","GA","Hawaii","HI","Idaho","ID","Illinois","IL","Indiana","IN","Iowa","IA","Kansas","KS","Kentucky","KY","Louisiana","LA","Maine","ME","Maryland","MD","Massachusetts","MA","Michigan","MI","Minnesota","MN","Mississippi","MS","Missouri","MO","Montana","MT","Nebraska","NE","Nevada","NV","New Hampshire","NH","New Jersey","NJ","New Mexico","NM","New York","NY","North Carolina","NC"," North Dakota","ND","Ohio","OH","Oklahoma","OK","Oregon","OR","Pennsylvania","PA","Rhode Island","RI","South Carolina","SC","South Dakota","SD","Tennessee","TN","Texas","TX","Utah","UT","Vermont","VT","Virginia","VA","Washington","WA","West Virginia","WV","Wisconsin","WI","Wyoming","WY"), byrow = TRUE, ncol = 2)

stateNames <- stateCodeMatrix[,1]
names(stateNames) <- stateCodeMatrix[,2]

stateAbbrevs <- stateAbbrevMatrix[,2]
names(stateAbbrevs) <- stateAbbrevMatrix[,1]

return(stateAbbrevs[stateNames[stateCode]])

}

roadFatalaties <- read.csv(file = "./../data/FARS/BikePedDeaths99-13.csv", na.strings = c("998","999"))
roadFatalaties <- within(roadFatalaties, {
  ageClass <- convertToAgeClass(AGE)
  SEX <- ifelse(SEX == 1, "M", ifelse(SEX == 2, "F", NA))
  SEX <- factor(SEX, levels = c("F","M"))
  STATE <- as.factor(convertState(STATE))
})
roadFatalaties <- subset(roadFatalaties, !is.na(SEX) ) #& STATE %in% c("California", "New York"))

p <- ggplot(roadFatalaties, aes(Type, fill = STATE))
p + geom_bar(position = "dodge") + theme_bw()+ theme(axis.text.x  = element_text(angle=90, vjust=0.5))# + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) #+ facet_grid( Type ~ ., scales = "free_y")

walkTime <- unlist(lapply(meanWalkList, function(x) x[3,2]))

foo <- data.frame(with(roadFatalaties, table(STATE,Type)))
within(subset(foo, Type == "Ped"), Walk <- walkTime[as.character(STATE)])

### 2016-10-05 manuscript/manuscript.tex

\begin{figure}[t]
    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig1}}
    \caption{Map of HHS Regions and metropolitan areas}\label{HHSmap}
\end{figure}

\section{U.S. Metro-Regional Analysis}

@@

\begin{figure}[t]
  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig3.pdf}}
    \caption{Mean travel time by mode (walking/cycling) for each of
      the four regions with a decrease in DALYs
      $(\textrm{HHS01},\textrm{HHS02},\textrm{HHS09},\textrm{HHS10})$. The
      national means are displayed on the far right.  These four
      regions correspond to New England/NewYork and the West
      Coast.}\label{meanMatrices}
\end{figure}

>>>>

\section{Seven State Analysis}

@@

\begin{figure}[t]
  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig3.pdf}}
    \caption{Mean travel time by mode (walking/cycling) for each of
      the four regions with a decrease in DALYs
      $(\textrm{HHS01},\textrm{HHS02},\textrm{HHS09},\textrm{HHS10})$. The
      national means are displayed on the far right.  These four
      regions correspond to New England/NewYork and the West
      Coast. The age classes are defined as in the original ITHIM
      implementation using 8 eight age classes defined in terms of
      years as follows: 1. 0--4, 2. 5--14, 3. 15--29, 4. 30--44,
      5. 45--59, 6. 60--69, 7. 70--79, and 8. 80+. Age class 1 is
      omitted from all calclulations in this paper due to lack of
      data.}\label{meanMatrices}
\end{figure}

\begin{figure}[t]
  \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig4.pdf}}
    \caption{California \& New York Averted DALYs Nationwide}\label{California}
\end{figure}

\begin{figure}[t]
    \centerline{\includegraphics[width=0.5\textwidth]{./figures/fig5.pdf}}
    \caption{Counts}\label{counts}
\end{figure}
